When the robot considers the human...

### 2013-04-23 paper.tex
Enriched the 'LAAS deliberative layer' part

\section{A Decisional framework}\label{sec:Framework}

@@

Our robot is controlled by a three layer
architecture~\cite{Alami1998}.  We present briefly its decisional
layer. The proposed decisional framework consists of several entities,
having each a specific role as illustrated by
Figure~\ref{architecture_fg}.  We describe how the robot is controlled
through an analysis of the three main activities performed by the
robot controller:

\begin{figure*}[thpb]
  \centering
  \includegraphics[width=1.0\textwidth]{./figs/archi.pdf} \\
  \caption {Architecture of the robot control system}
  \label{architecture_fg}
\end{figure*}

@@

The next three sections describe the three robot controller
activities and how they make use of a number of key components in the
architecture:

\begin{itemize}
\item SPARK: Spatial Reasoning and Knowledge module \cite{Sisbot2011}
\item ORO: a knowledge management module \cite{Lemaignan2010}
\item HATP: a Human-Aware Task Planner \cite{Alili2008}
\item A set of Human aware motion, placement and manipulation planners
 \cite{Sisbot2008, Mainprice2011, Pandey2010}
\end{itemize}

>>>>

\section{The LAAS deliberative layer}
\label{sec:Framework}

Fig.~\ref{fig|archi} give an overview of the connections between the
deliberative components of our architecture~\cite{Alami2011}.

\begin{figure*}
        \centering
        \includegraphics[width=1.0\columnwidth]{figs/archi}
        \caption{Overview of the LAAS deliberative layer. Knowledge is
        centrally managed in an active \emph{semantic blackboard}, pictured
        above with a thick border.}
        \label{fig|archi}
\end{figure*}

This architecture moves away from standard layered approaches. Interactions
between components are mostly bidirectional and, from the software components
point of view, we do not introduce layers of abstraction (we do, however, have
access to the lower level modules of the robot to execute
actions~\cite{Alami1998}, but all cognition-related modules reside at the same
level). This is especially visible for the dialogue input processing. This
component does not simply act as an alternative perceptual input to the
symbolic database, but also actively queries previously acquired knowledge to
disambiguate and validate the newly created symbolic knowledge (see
section~\ref{sect|com}).

Our architecture relates but is to be distinguished from \emph{Beliefs,
Desires, Intentions} (BDI) architectures. BDI architectures are primarily
focused on \emph{practical reasoning}, \ie the process of deciding, step by
step, which action to perform to reach a goal (as summarised by
Woolridge~\cite{Woolridge1999}). The management of the interaction between
knowledge (the beliefs) and task and plan representation and execution (the
desires and the intentions) is central, and aims at selecting at each step the
best subgoal. It becomes then an intention that the robot commits to.

This interaction between knowledge and actions is also central to our approach
(as for any cognitive system), but task representation and task execution is
not seen as a monolithic, central function: it is one of the activities of the
robot, actually split between communication components (that can acquire
desires from interaction with agents, amongst other things) and an execution
controller that may decide to take an incoming desire into account to create
its own internal goals. The controller generates and controls intentions from
these goals with the help of a symbolic task planner, that has also direct
access to the knowledge base.

The architecture is not only focused on this workload, and other
activities are conducted without being explicitly considered as desires:
assessment of the situation and the environment, dialogue (including
performative dialogue that can possibly change the internal state of the robot,
but does not lead to the creation of desires,  like question answering or
statement assertion), various background monitoring and recognition tasks, etc.

Regarding the anchoring question, this architecture is bidirectional. The
components we described provide a \textit{bottom-up} grounding process:
geometric reasoning and dialogue processing modules constantly build and push
new symbolic contents about the world to the knowledge base where it becomes
accessible to decisional layers. In parallel, the knowledge base relies on
reasoning in a \textit{top-down} way to produce new facts that may trigger in
return physical behaviours.

\subsection*{Knowledge model}

In our architecture (Fig.~\ref{fig|archi}), knowledge manipulation relies on a
\emph{semantic blackboard}: one central server (the {\sc Oro}
server~\cite{Lemaignan2010}) stores knowledge as it is produced by each of the
deliberative components. It conversely exposes an API to query the knowledge
base.

Knowledge is represented as RDF triples in the OWL sub-language. Each
time triples are added or removed from the knowledge base, a Description
Logics reasoner ({\sc Pellet}\footnote{\url{http://clarkparsia.com/pellet/}})
classifies the whole ontology and inserts all possible inferred triples.

Relying on RDF triples and Description Logics has major advantages like the
availability of numerous open-source libraries to manipulate the ontology,
interoperability with several major on-line knowledge bases (like {\sc
OpenCyc}, {\sc WordNet} or {\sc DBPedia}), open-world reasoning, and the formal
guarantee of decidability (it is always possible to classify a Description
Logics ontology).

It has also notable limitations: RDF triples imply only binary predicates
(\stmt{subject predicate object}), which constrains the expressiveness of the
system or leads to cumbersome reifications. Alternative exists (like {\sc
KnowRob}~\cite{Tenorth2009a}) that mixes RDF with more expressive logic
languages like {\sc Prolog}, at the price, however, of other limitations, like
closed-world reasoning or immutable T-Box. The classification
performance is another issue: from our experience, with an ontology sized for a
standard experiment (about 100 classes and 200 instances), classification
typically take about 100ms, which becomes problematic during interactions.
Besides, the performances are difficult to predict, since a seemingly
inoffensive new statement may indirectly change radically the logical
complexity of the whole knowledge model and lead to notable degradation of
classification time.

This knowledge model also largely exclude representation of continuous
phenomena (like time) or uncertain phenomena. When required, these have to be
managed by dedicated components.

Amongst the features of {\sc Oro}, the system can dynamically creates
new independent knowledge models (\ie independent instantiations of the
ontology). This feature is used for instance to store separately the beliefs of
each of the agents the robot interacts with (see section~\ref{sect|tom}).

\subsection*{Interaction model}

@@

Next sections describe how the robot is controlled through an analysis of the
three main activities performed by the robot controller:

@@

These activities are caried out by different software composants:

\begin{itemize}

    \item Situation assessment is managed by \textsc{Spark} (Spatial Reasoning
        and Knowledge module)~\cite{Sisbot2011},

    \item HATP (Human-Aware Task Planner)~\cite{Alili2008} is responsible for
        symbolic task planning,

    \item We finally also rely on a set of human-aware motion, placement and
        manipulation planners, grouped under the \textsc{Move3D}
        umbrella~\cite{Sisbot2008, Mainprice2011, Pandey2010}.

\end{itemize}
